Measurement of nuclear excitation functions for proton induced reactions (E\texorpdfstring{$_{\text{p}}$\,=\,}{Ep = 40--90}40-90 MeV) on natural Nb

### 2018-02-23 nb_px_paper/nb_px_paper.tex
Corrected hyphen/en-dash/em-dash usage. Squashed various comments.

A stacked target of natural Nb, Cu, and Al thin foils were irradiated by 100 MeV protons at the Los Alamos National Laboratory's Isotope Production Facility.
The irradiation was motivated by interest in the characterization of the \ce{^{93}Nb}(p,4n)\ce{^{90}Mo} reaction as a new dosimetry standard for intermediate-energy protons.
A set of 38 measurements of cross sections for the \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions between 40--90 MeV are presented in this work, as well as 5 independent measurements of isomer branching ratios.
Cross sections were measured using the well-characterized 
% \ce{^{nat}Al}(p,x)\ce{^{22}Na}, \ce{^{nat}Al}(p,x)\ce{^{24}Na}, 
\ce{^{nat}Cu}(p,x)\ce{^{56}Co}, \ce{^{nat}Cu}(p,x)\ce{^{62}Zn}, and \ce{^{nat}Cu}(p,x)\ce{^{65}Zn} monitor reactions  for accurate proton fluence measurement, with all activities quantified using high-purity germanium spectroscopy.
Variance minimization techniques were employed to reduce systematic uncertainties in proton energy and fluence, improving the reliability of these measurements. 
\textred{The \ce{^{nat}Al}(p,x)\ce{^{22}Na}, \ce{^{nat}Al}(p,x)\ce{^{24}Na}, }
The measured cross sections are shown to be in excellent agreement with  literature values, and have been measured with improved precision compared with previous measurements.
This work reports the first measurements of the  \ce{^{nat}Nb}(p,x)\ce{^{82m}Rb} reaction, as well as the first measurement of the independent cross sections for    
% \ce{^{nat}Cu}(p,x)\ce{^{52\text{m}}Mn}, 
\ce{^{nat}Cu}(p,x)\ce{^{52\text{g}}Mn} and \ce{^{nat}Nb}(p,x)\ce{^{85\text{g}}Y} in the 40--90 MeV region.
Comparisons against the reaction modeling codes CoH$_3$, EMPIRE, and TALYS are used to explore the impact of pre-equilibrium particle emission in the reaction dynamics for 40--90 MeV \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions.

@@

Every year, approximately 17 million nuclear medicine procedures (both diagnostic and therapeutic) are performed in the U.S. alone,  which has made incredible strides in improving our ability to detect and treat a variety of life-threatening diseases \cite{Delbeke2011,NSACIsotopesSubcommittee2015}.
The vast majority of the radioisotopes currently used for these procedures are produced in the field's array of low- (E \textless\ 30 MeV / A) and intermediate-energy (30 \textless\ E \textless\ 200 MeV / A) accelerator capabilities, which routinely produce many of the staple medical radionuclides, such as \ce{^{11}C}, \ce{^{18}F}, \ce{^{68}Ge}, \ce{^{82}Rb}, and \ce{^{123}I}, as well as many of the non-medical radioisotopes of commercial value, such as  \ce{^{32}Si}, \ce{^{73}As}, \ce{^{95m}Tc}, and  \ce{^{109}Cd} \cite{international2009iaea,schlyer2008cyclotron}. 
\comment{Should \enquote{intermediate-energy} be less than 200 MeV, or 100 MeV?  I find conflicting descriptions in the literature.\\Lee:  I believe that 60 MeV is considered as the transition point\\ Stephen: In this context, I would make this definition based on what machines in the US are routinely used for medical isotope production - based on this, I agree with 200 MeV. }
 The future of nuclear medicine would appear to be the paradigm of personalized medicine --- targeted radionuclide therapy to spare healthy tissue \cite{Mulford2005,Qaim201731}, and theranostic medicine, which pairs an imaging isotope with a therapeutic isotope (frequently, of the same element), to provide simultaneous, real-time dose delivery and verification, leading to drastic reductions in prescribed patient dose \cite{Muller2014,Bentzen2005,Srivastava2012}. 
Other variants of theranostic medicine exist, including pre-imaging for treatment planning, or delivery of a single compound with different radioelements for imaging/therapy where the inter-element biodistribution has been validated.  
Candidate isotopes to meet these needs have been identified based on their chemical and radioactive decay properties \cite{NSACIsotopesSubcommittee2015,Qaim201731,bernstein2015nuclear}, and a series of campaigns are underway to perform targeted, high-priority measurements of thin-target cross sections and thick-target integral yields.
These studies will serve to facilitate the production of pre-clinical quantities of radioactivity.

One significant obstacle to both high-fidelity measurements of production cross sections for emerging medical radionuclides, as well as conventional isotope production, is a lack of well-characterized  dosimetry standards.  
This is particularly true for higher-energy charged particle beams, where there is currently a paucity of such data. 
Indeed, the development of new dosimetry standards and the improved evaluation of existing standards is one of the areas of greatest cross-cutting need for nuclear data \cite{bernstein2015nuclear}. 
Charged particle dosimetry data currently exists for several low-to-intermediate energy charged particle beams (E \textless\ 50 MeV / A), but even the experimental data used for this evaluation are  sparse above approximately 30 MeV / A and  uncertainties in experimental cross sections are large (6-15\%) \cite{gul2001charged}. 
While work is needed to improve upon existing dosimetry data, the development of new reactions can expand the available range of options for the monitoring of charged particle beams.
\comment{Stephen: I'm having a hard time following this paragraph. Some of your syntax overlaps significantly with nuclear medicine \& medical physics terminology with different intended meanings. For example 'dosimetry' is typically taken to mean the measurement of absorbed energy per unit mass (J/kg) and associated measurements/specifications. Another example, 'dosimetry standard' is used quite often when referring to NIST air kerma reference standards which are the bedrock of Accredited Dosimetry Calibration Lab (ADCL) measurements. As you are more familiar with classic Nuclear Physics writing, perhaps you better know how this language will be interpreted. }

@@

Moving to the higher energy  of the reaction's pre-equilibrium tail, the excitation function becomes  smooth and generally flat as a function of energy.
In this regime, the monitor reaction offers little-to-no energy sensitivity. 
However,  in the pre-equilibrium regime, monitor reactions become extremely useful for determining the integral beam current. 
While cross section measurements often use external beam current monitors (such as an inductive pickup upstream of a target, or an electrically-isolated target in a Faraday cup), these measure the integrated current incident upon an entire target assembly.
For the case of stacked-target activation experiments, commonly employed to measure cross sections at multiple energies  in a single activation, external beam current monitors can only measure the integral current incident upon the \enquote{front} (upstream) of the target stack.
In these experiments, a series of monitor foils at each energy position allow one to measure the integral current at each position in the stack, reducing systematic errors in observed cross section magnitude.
\comment{Stephen:  ``and precision, unfortunately ''.  Update post-discussion.}
Both of these purposes make well-characterized dosimetry an invaluable asset to any activation experiment.

@@

From a targetry  perspective, it is preferable to use a naturally mono-isotopic target that is readily commercially available at an affordable price and is generally chemically inert - any significant chemical changes during target preparation (significant oxidation, etc) will affect the target's areal density, systematically changing the measured integral current. 
Structurally, the target material should be malleable and supportive to be able to be formed into a thin target.
For charged particle reactions,  energy degradation scales with target areal density,  broadening the energy spectrum downstream of the target.
However, since the monitor reaction yield also scales with target areal density, the use of a target which is too thin may provide insufficient counting statistics during decay spectroscopy.
For reference, a monitor foil of approximately 25 mg/cm$^2$ provides a good compromise, with less than 100 keV degradation for a proton energy of 100 MeV, and less than 200 keV at 40 MeV.
Thickness selection will be subject to the context of an experiment, seeking to maximize thickness without overly perturbing the energy uncertainty of  measurements.
% These are the primary characteristics involved in choosing an appropriate target for a monitor reaction

\comment{Stephen:  This is also a discussion point. If you're trying to introduce the purpose of this paper, i.e. the establishment of a new monitor reaction, I would try to condense paragraphs 3-9 into two paragraphs: one paragraph explaining what a monitor reaction is used for, and the other describing considerations relevant to the use of a monitor reaction. The rest of this content can be included in the discussion if you feel it is necessary.}
Lastly, and perhaps most importantly for high-energy dosimetry applications, it is  of utmost importance to choose a reaction channel which cannot be populated via secondary particles incident upon the monitor target.
This is typically mostly a concern for secondary neutrons produced through (z,xn) reactions on upstream targets, degraders, and stack materials, to avoid monitor reactions which can be populated through (n,x) reactions on the target.
Any monitor reaction channel which can be populated by anything other than the primary beam should be avoided, as it is often a laborious task to separate out the fraction of secondary particles contributing to the total activation.

% The purpose of the present work is to explore the potential to use high-flux neutron generators to produce high-specific activity samples of radionuclides at the mCi level for local use in the application community. 
%  The research group at UC Berkeley has  developed a High Flux Neutron Generator (HFNG) that features an internal target where samples can be placed just several millimeters from the neutron producing surface in order to maximize the utilization of the neutron yield for the production of a desired radionuclide \cite{Waltz2017,Waltz2016a,doi:10.1063/1.3267832}.
%  The HFNG uses the D(D,n)\ce{^3He} reaction to produce neutrons with energies near 2.45 MeV together with a self-loading target design to maintain continuous operation without target replacement.
%  In addition to the generator itself, efforts are underway to design neutron reflection capabilities to allow scattered neutrons multiple opportunities to interact with an  internally mounted target.
% While these design efforts are underway, the HFNG can be used to better characterize production cross sections at the appropriate neutron energy.
% 
% 
% The present work features a pair of cross section measurements for the production of two emerging non-standard medical radionuclides: the positron emitter \ce{^{64}Zn}(n,p)\ce{^{64}Cu} and the single - photon emission computed tomography (SPECT) tracer \ce{^{47}Ti}(n,p)\ce{^{47}Sc}.
% \ce{^{64}Cu}  ($t_{1/2}$ = 12.7 h) undergoes $\beta^+$ decay (61.5\% branching ratio) to \ce{^{64}Ni} or $\beta^-$ decay (38.5\% branching ratio) to \ce{^{64}Zn} \cite{Singh2007}.
% The emitted short-range 190-keV $\beta^-$ particle makes this an  attractive  therapeutic radionuclide, which also has the possibility for simultaneous positron emission tomography (PET) imaging for real-time dose monitoring and verification.
% This makes \ce{^{64}Cu} particularly desirable  for emerging radiation therapy protocols \cite{Lewis2003,NSACIsotopesSubcommittee2015,Bandari2014,mp500671j}.
% In addition, copper radiochemistry is well developed, and many existing ligands and carriers may be used for selective delivery of the radionuclide to different sites in patients.
% The second radionuclide studied, \ce{^{47}Sc} ($t_{1/2}$ = 3.35 d), undergoes $\beta^-$ decay to \ce{^{47}Ti}, emitting a high-intensity (63.8\%) 159-keV gamma-ray in the process \cite{Burrows2007}.
% This radionuclide is  attractive as an emerging diagnostic isotope, due to the similarity of the emitted gamma-ray to that of the  well-established \ce{^{99m}Tc} \cite{Qaim2011,Qaim201731,Kolsky1998,mausner1995evaluation}.
% Due to the short half-life ($t_{1/2}$ = 6.0 h) of and dwindling supplies of \ce{^{99m}Tc}, \ce{^{47}Sc} stands poised as a potential solution to this shortage, due to its longer half-life and multiple production pathways without the need for highly enriched uranium \cite{Browne2011}.
% In addition, when paired with \ce{^{44}Sc}, \ce{^{47}Sc} forms a promising \enquote{theranostic} pair for use in simultaneous therapeutic and diagnostic applications \cite{Muller2014,Deilami-nezhad2016}.

One monitor reaction which satisfies these requirements is that of a new, intermediate-energy proton dosimetry standard based on \ce{^{93}Nb}(p,4n)\ce{^{90}Mo}. 
Niobium is naturally mono-isotopic, readily  available commercially in high purity, is chemically inert, and can easily be rolled down to foils as thin as 1 $\micro$m.  
\ce{^{90}Mo} also has excellent decay properties - its fairly long-lived lifetime ($\epsilon=100\%, t_{1/2}=5.56 \pm 0.09$ h) allows it to be counted at leisure without fear of the product \ce{^{90}Mo} decaying away excessively between end-of-beam and the start of counting, and it possesses seven strong, distinct gamma lines (notably its 122.370 keV ($I_\gamma = 64 \pm 3\%$) and 257.34 keV ($I_\gamma = 78 \pm 4\%$) lines) which can be used to uniquely and easily   quantify \ce{^{90}Mo} production \cite{Browne1997}. 
In addition,  \ce{^{90}Mo} is completely immune from (n,x) production on  \ce{^{93}Nb}, being produced only via the primary proton beam, and the \ce{^{90}Mo} decay lines can only be observed in its decay, as its daughter, \ce{^{90}Nb}, is also unstable and decays via $\epsilon$ to stable \ce{^{90}Zr}.

The purpose of the present work is to  measure the production of the long-lived radionuclide \ce{^{90}Mo} ($t_{1/2}=5.56 \pm 0.09$ h \cite{Browne1997}) via the $^\text{nat}$Nb(p,x) reaction. 
In addition to the $^\text{nat}$Nb(p,x)\ce{^{90}Mo} measurement, this experiment has also yielded measurements of 32 other (p,x) production cross sections between 40 -- 90 MeV  for a number of additional reaction products, including several emerging radionuclides with medical applications.
These include the non-standard positron emission tomography (PET) agent \ce{^{57}Ni}  ($t_{1/2}=35.60\pm0.06$ h \cite{Bhat1998}), \ce{^{64}Cu} ($t_{1/2}=12.701 \pm 0.002$ h \cite{Singh2007}), \ce{^{86}Y}  ($t_{1/2}=14.74\pm0.02$ h \cite{NEGRET20151}), \ce{^{89}Zr} ($t_{1/2}=78.41\pm0.12$ h \cite{Singh2013}),  \ce{^{90}Nb} ($t_{1/2}=14.60 \pm 0.05$ h \cite{Browne1997}),  and the diagnostic agent \ce{^{82\text{m}}Rb} ($t_{1/2}=6.472\pm0.006$ h \cite{Tuli2003}).

@@

A stacked-target design was utilized for this work in order that the (p,x) cross sections for each reaction channel could be measured at multiple energy positions in a single irradiation \cite{Cumming1963}. 
For targets, a series of nominal 25 \micro m \ce{^{nat}Nb} foils (99.8\%, lot \#T23A035), 25 \micro m \ce{^{nat}Al} foils (99.999\%, lot \#M06C032), and 50 \micro m \ce{^{nat}Cu} foils (99.9999\%, lot \#N26B062) were used (all from Alfa Aesar, Ward Hill, MA, 01835, USA).
Six foils of each metal were cut down to 2.5 x 2.5 cm squares and characterized - for each foil, length and width measurements were taken at four different locations using a digital caliper (Mitutoyo America Corp.), thickness measurements were taken at four different locations using a digital micrometer (Mitutoyo America Corp.), and four mass measurements were taken using an analytical balance after cleaning the foils with isopropyl alcohol.
Using these length, width, and mass readings, the areal density (in mg/cm$^2$) for each foil was calculated, along with the propagated uncertainty in areal density.
The foils were tightly sealed into \enquote{packets} using two pieces of  3M 5413-Series Kapton polyimide film tape - each piece of tape consists of 43.2 \micro m of a silicone adhesive (nominal 4.793 mg/cm$^2$) on 25.4 \micro m of a polyimide backing (nominal 3.607 mg/cm$^2$).
The sealed foils were mounted over the hollow center of a 1.575 mm-thick plastic frame.
One \ce{^{nat}Al}, one \ce{^{nat}Cu}, and \ce{^{nat}Nb} mounted foil were bundled together using baling wire for each energy position.
These foil packet bundles were lowered into the beamline by inserting them into a  water-cooled production target box.
% These six foil packet bundles were inserted into one of six  
The box, seen in \autoref{fig:target_stack}, is machined from 6061 aluminum alloy, has a thin (0.64 mm) Inconel beam entrance window, and  contains 6 \enquote{energy positions} for targets, formed by  5 slabs of 6061 aluminum alloy (previously characterized) which serve as proton energy degraders  between energy positions.
At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor - after end-of-beam (EoB), decay radiation emitted from these activated stainless steel foils may be used to develop radiochromic film (Gafchromic EBT3), revealing the spatial profile of the beam entering and exiting the stack.
After loading all targets in the stack, the lid of the target box is sealed in place, using an inset o-ring to create a water-tight seal, and the box is lowered through a hot cell into the beamline, where it sits electrically isolated.
The specifications of the target stack design for this work is presented in \autoref{tab:stack_table}.

This target stack was assembled and irradiated at the Isotope Production Facility (IPF) at the Los Alamos National Laboratory (LANL), using the LANSCE linear accelerator. 
The stack was irradiated for approximately 2 hours with a nominal current of 1 mA, using a 50 \micro s pulse at a frequency of 2 Hz, for an anticipated integral current of 205.9 nAh.
The beam current, measured using an inductive pickup, remained stable under these conditions for the duration of the irradiation, with the exception of approximately 70\,s of downtime, which occurred approximately 3 minutes into irradiation.
The proton beam incident upon the stack's Inconel beam entrance window had an average energy of 100 MeV determined via time-of-flight, with an approximately Gaussian energy distribution width of 0.1 MeV - this energy profile was used for all later analysis.
At the end of the irradiation, the target stack was withdrawn from the beamline into the IPF hot cell, where it was disassembled and the activated foils removed using robotic manipulators.
The activated foils were cleaned of all surface contamination, and transported to a counting lab for gamma spectroscopy, which started approximately 7 hours following end-of-beam.

@@

For consistency, a single detector was used in this measurement, an ORTEC GEM Series (model \#GEM10P4-70)  High-Purity Germanium (HPGe) detector.
The detector is a mechanically-cooled coaxial p-type HPGe with a 1 mm aluminum window, and a 49.2 mm diameter, 27.9 mm long crystal.
Samples were counted at fixed positions ranging 4.5 - 83.5  cm (5\% maximum permissible dead-time) from the front face of the detector, with a series of standard calibration sources used to determine energy, efficiency, and pileup calibrations for each position.
The foils were counted  for a period of 2 weeks following EoB, to accurately quantify all induced activities,  with dead time never exceeding 5\%.
An example gamma-ray spectrum collected in such a fashion is shown in \autoref{fig:gspec}.
For all spectra collected, net peak areas were fitted using the gamma spectroscopy analysis code UNISAMPO \cite{Aarnio2001}, which has been shown to perform best in comparisons with other common analysis codes \cite{Jackman2014}.
%, due to its peak fitting algorithms and incorporation  of calibrated detector-specific peak shapes and widths.

@@

The \enquote{variance minimization} techniques  described by  Graves \etal\  have been employed here to further reduce the uncertainty in proton energy assignments     \cite{Graves2016}.
This method is based on the assumption that the independent measurements of proton fluence from the 5 monitor reactions used in this work should all be consistent at each energy position.
If the monitor reaction cross sections and MCNP6 - modeled energy distributions are both accurate, then any disagreement in the  observed proton fluences is due to a systematic uncertainty in the stack design, namely, the areal densities of the stack components \cite{Graves2016,Marus2015}. 
This disagreement is minor at the front of the stack, and gets progressively worse as the beam is degraded, due to the compounded effect of systematic uncertainties in stack areal densities.

>>>>

A stacked target of natural Nb, Cu, and Al thin foils were irradiated by 100 MeV protons at the Los Alamos National Laboratory's Isotope Production Facility.
The irradiation was motivated by interest in the characterization of the \ce{^{93}Nb}(p,4n)\ce{^{90}Mo} reaction as a new dosimetry standard for intermediate-energy protons.
A set of 38 measurements of cross sections for the \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions between 40--90 MeV are presented in this work, as well as 5 independent measurements of isomer branching ratios.
Cross sections were measured using the well-characterized 
% \ce{^{nat}Al}(p,x)\ce{^{22}Na}, \ce{^{nat}Al}(p,x)\ce{^{24}Na}, 
\ce{^{nat}Cu}(p,x)\ce{^{56}Co}, \ce{^{nat}Cu}(p,x)\ce{^{62}Zn}, and \ce{^{nat}Cu}(p,x)\ce{^{65}Zn} monitor reactions  for accurate proton fluence determination, with all activities quantified using high-purity germanium spectroscopy.
Variance minimization techniques were employed to reduce systematic uncertainties in proton energy and fluence, improving the reliability of these measurements. 
\textred{The \ce{^{nat}Al}(p,x)\ce{^{22}Na} and \ce{^{nat}Al}(p,x)\ce{^{24}Na}  monitor reactions  were excluded from proton fluence determination due to the production of \ce{^{22,24}Na} contamination by \ce{^{nat}Si}(p,x) activation of the silicone adhesive used for sealing foil packets.}}
The measured cross sections are shown to be in excellent agreement with  literature values, and have been measured with improved precision compared with previous measurements.
This work reports the first measurements of the  \ce{^{nat}Nb}(p,x)\ce{^{82m}Rb} reaction, as well as the first measurement of the independent cross sections for    
% \ce{^{nat}Cu}(p,x)\ce{^{52\text{m}}Mn}, 
\ce{^{nat}Cu}(p,x)\ce{^{52\text{g}}Mn} and \ce{^{nat}Nb}(p,x)\ce{^{85\text{g}}Y} in the 40--90 MeV region.
Comparisons against the reaction modeling codes CoH$_3$, EMPIRE, and TALYS are used to explore the impact of pre-equilibrium particle emission in the reaction dynamics for 40--90 MeV \ce{^{nat}Nb}(p,x) and  \ce{^{nat}Cu}(p,x) reactions.

@@

Every year, approximately 17 million nuclear medicine procedures (both diagnostic and therapeutic) are performed in the U.S. alone,  which has made incredible strides in improving our ability to detect and treat a variety of life-threatening diseases \cite{Delbeke2011,NSACIsotopesSubcommittee2015}.
The vast majority of the radioisotopes currently used for these procedures are produced in the field's array of low- (E \textless\ 30 MeV / A) and intermediate-energy (30 \textless\ E \textless\ 200 MeV / A) accelerator capabilities, which routinely produce many of the staple medical radionuclides, such as \ce{^{11}C}, \ce{^{18}F}, \ce{^{68}Ge}, \ce{^{82}Rb}, and \ce{^{123}I}, as well as many  
% of the 
non-medical radioisotopes of commercial value, such as  \ce{^{32}Si}, \ce{^{73}As}, \ce{^{95m}Tc}, and  \ce{^{109}Cd} \cite{international2009iaea,schlyer2008cyclotron}. 
\comment{Should \enquote{intermediate-energy} be less than 200 MeV, or 100 MeV?  I find conflicting descriptions in the literature.\\Lee:  I believe that 60 MeV is considered as the transition point\\ Stephen: In this context, I would make this definition based on what machines in the US are routinely used for medical isotope production - based on this, I agree with 200 MeV. }
The future of nuclear medicine would appear to be the paradigm of personalized medicine --- targeted radionuclide therapy to spare healthy tissue \cite{Mulford2005,Qaim201731}, and theranostic medicine, which pairs an imaging isotope with a therapeutic isotope (frequently, of the same element), to provide simultaneous, real-time dose delivery and verification, leading to drastic reductions in prescribed patient dose \cite{Muller2014,Bentzen2005,Srivastava2012}. 
Other variants of theranostic medicine exist, including pre-imaging for treatment planning, or delivery of a single compound with different radioelements for imaging/therapy where the inter-element biodistribution has been validated.  
Candidate isotopes to meet these needs have been identified based on their chemical and radioactive decay properties \cite{NSACIsotopesSubcommittee2015,Qaim201731,bernstein2015nuclear}, and a series of campaigns are underway to perform targeted, high-priority measurements of thin-target cross sections and thick-target integral yields.
These studies will serve to facilitate the production of pre-clinical quantities of radioactivity.

One significant obstacle to both high-fidelity measurements of production cross sections for emerging medical radionuclides, as well as conventional isotope production, is a lack of well-characterized  dosimetry standards.  
This is particularly true for higher-energy charged particle beams, where there is currently a paucity of such data. 
Indeed, the development of new dosimetry standards and the improved evaluation of existing standards is one of the areas of greatest cross-cutting need for nuclear data \cite{bernstein2015nuclear}. 
Charged particle dosimetry data currently exists for several low-to-intermediate energy charged particle beams (E \textless\ 50 MeV / A), but even the experimental data used for this evaluation are  sparse above approximately 30 MeV / A and  uncertainties in experimental cross sections are large (6--15\%) \cite{gul2001charged}. 
While work is needed to improve upon existing dosimetry data, the development of new reactions can expand the available range of options for the monitoring of charged particle beams.
\comment{Stephen: I'm having a hard time following this paragraph. Some of your syntax overlaps significantly with nuclear medicine \& medical physics terminology with different intended meanings. For example 'dosimetry' is typically taken to mean the measurement of absorbed energy per unit mass (J/kg) and associated measurements/specifications. Another example, 'dosimetry standard' is used quite often when referring to NIST air kerma reference standards which are the bedrock of Accredited Dosimetry Calibration Lab (ADCL) measurements. As you are more familiar with classic Nuclear Physics writing, perhaps you better know how this language will be interpreted. }

@@

Moving to the higher energy  of the reaction's pre-equilibrium tail, the excitation function becomes  smooth and generally flat as a function of energy.
In this regime, the monitor reaction offers little-to-no energy sensitivity. 
However,  in the pre-equilibrium regime, monitor reactions become extremely useful for determining the integral beam current. 
While cross section measurements often use external beam current monitors (such as an inductive pickup upstream of a target, or an electrically-isolated target in a Faraday cup), these measure the integrated current incident upon an entire target assembly.
For the case of stacked-target activation experiments, commonly employed to measure cross sections at multiple energies  in a single activation, external beam current monitors can only measure the integral current incident upon the \enquote{front} (upstream) of the target stack.
In these experiments, a series of monitor foils at each energy position allows one to indirectly measure the integral current at each position in the stack, reducing systematic errors in observed cross section magnitude, but with reduced precision compared to direct measurement using a well-characterized suppressed Faraday cup.
% \comment{Stephen:  ``and precision, unfortunately ''.  Update post-discussion.}
Both of these purposes make well-characterized dosimetry an invaluable asset to any activation experiment.

@@

From a targetry  perspective, it is preferable to use a naturally mono-isotopic target that is readily commercially available at an affordable price and is generally chemically inert --- any significant chemical changes during target preparation (significant oxidation, etc) will affect the target's areal density, systematically changing the measured integral current. 
Structurally, the target material should be malleable and supportive to be able to be formed into a thin target.
For charged particle reactions,  energy degradation scales with target areal density,  broadening the energy spectrum downstream of the target.
However, since the monitor reaction yield also scales with target areal density, the use of a target which is too thin may provide insufficient counting statistics during decay spectroscopy.
For reference, a monitor foil of approximately 25 mg/cm$^2$ provides a good compromise, with less than 100 keV degradation for a proton energy of 100 MeV, and less than 200 keV at 40 MeV.
Thickness selection will be subject to the context of an experiment, seeking to maximize thickness without overly perturbing the energy uncertainty of  measurements.
% These are the primary characteristics involved in choosing an appropriate target for a monitor reaction

\comment{Stephen:  This is also a discussion point. If you're trying to introduce the purpose of this paper, i.e. the establishment of a new monitor reaction, I would try to condense paragraphs 3--9 into two paragraphs: one paragraph explaining what a monitor reaction is used for, and the other describing considerations relevant to the use of a monitor reaction. The rest of this content can be included in the discussion if you feel it is necessary.}
Lastly, and perhaps most importantly for high-energy dosimetry applications, it is  of utmost importance to choose a reaction channel which cannot be populated via secondary particles incident upon the monitor target.
This is typically mostly a concern for secondary neutrons produced through (z,xn) reactions on upstream targets, degraders, and stack materials, to avoid monitor reactions which can be populated through (n,x) reactions on the target.
Any monitor reaction channel which can be populated by anything other than the primary beam should be avoided, as it is often a laborious task to separate out the fraction of secondary particles contributing to the total activation.

One monitor reaction which satisfies these requirements is that of a new, intermediate-energy proton dosimetry standard based on \ce{^{93}Nb}(p,4n)\ce{^{90}Mo}. 
Niobium is naturally mono-isotopic, readily  available commercially in high purity, is chemically inert, and can easily be rolled down to foils as thin as 1 $\micro$m.  
\ce{^{90}Mo} also has excellent decay properties --- its fairly long-lived lifetime ($\epsilon=100\%, t_{1/2}=5.56 \pm 0.09$ h) allows it to be counted at leisure without fear of the product \ce{^{90}Mo} decaying away excessively between end-of-beam and the start of counting, and it possesses seven strong, distinct gamma lines (notably its 122.370 keV ($I_\gamma = 64 \pm 3\%$) and 257.34 keV ($I_\gamma = 78 \pm 4\%$) lines) which can be used to uniquely and easily   quantify \ce{^{90}Mo} production \cite{Browne1997}. 
In addition,  \ce{^{90}Mo} is completely immune from (n,x) production on  \ce{^{93}Nb}, being produced only via the primary proton beam, and the \ce{^{90}Mo} decay lines can only be observed in its decay, as its daughter, \ce{^{90}Nb}, is also unstable and decays via $\epsilon$ to stable \ce{^{90}Zr}.

The purpose of the present work is to  measure the production of the long-lived radionuclide \ce{^{90}Mo} ($t_{1/2}=5.56 \pm 0.09$ h \cite{Browne1997}) via the $^\text{nat}$Nb(p,x) reaction. 
In addition to the $^\text{nat}$Nb(p,x)\ce{^{90}Mo} measurement, this experiment has also yielded measurements of 32 other (p,x) production cross sections between 40--90 MeV  for a number of additional reaction products, including several emerging radionuclides with medical applications.
These include the non-standard positron emission tomography (PET) agent \ce{^{57}Ni}  ($t_{1/2}=35.60\pm0.06$ h \cite{Bhat1998}), \ce{^{64}Cu} ($t_{1/2}=12.701 \pm 0.002$ h \cite{Singh2007}), \ce{^{86}Y}  ($t_{1/2}=14.74\pm0.02$ h \cite{NEGRET20151}), \ce{^{89}Zr} ($t_{1/2}=78.41\pm0.12$ h \cite{Singh2013}),  \ce{^{90}Nb} ($t_{1/2}=14.60 \pm 0.05$ h \cite{Browne1997}),  and the diagnostic agent \ce{^{82\text{m}}Rb} ($t_{1/2}=6.472\pm0.006$ h \cite{Tuli2003}).

@@

A stacked-target design was utilized for this work in order that the (p,x) cross sections for each reaction channel could be measured at multiple energy positions in a single irradiation \cite{Cumming1963}. 
For targets, a series of nominal 25 \micro m \ce{^{nat}Nb} foils (99.8\%, lot \#T23A035), 25 \micro m \ce{^{nat}Al} foils (99.999\%, lot \#M06C032), and 50 \micro m \ce{^{nat}Cu} foils (99.9999\%, lot \#N26B062) were used (all from Alfa Aesar, Ward Hill, MA, 01835, USA).
Six foils of each metal were cut down to 2.5 x 2.5 cm squares and characterized --- for each foil, length and width measurements were taken at four different locations using a digital caliper (Mitutoyo America Corp.), thickness measurements were taken at four different locations using a digital micrometer (Mitutoyo America Corp.), and four mass measurements were taken using an analytical balance after cleaning the foils with isopropyl alcohol.
Using these length, width, and mass readings, the areal density (in mg/cm$^2$) for each foil was calculated, along with the propagated uncertainty in areal density.
The foils were tightly sealed into \enquote{packets} using two pieces of  3M 5413-Series Kapton polyimide film tape --- each piece of tape consists of 43.2 \micro m of a silicone adhesive (nominal 4.79 mg/cm$^2$) on 25.4 \micro m of a polyimide backing (nominal 3.61 mg/cm$^2$).
The sealed foils were mounted over the hollow center of a 1.575 mm-thick plastic frame.
One \ce{^{nat}Al}, one \ce{^{nat}Cu}, and \ce{^{nat}Nb} mounted foil were bundled together using baling wire for each energy position.
These foil packet bundles were lowered into the beamline by inserting them into a  water-cooled production target box.
% These six foil packet bundles were inserted into one of six  
The box, seen in \autoref{fig:target_stack}, is machined from 6061 aluminum alloy, has a thin (0.64 mm) Inconel beam entrance window, and  contains 6 \enquote{energy positions} for targets, formed by  5 slabs of 6061 aluminum alloy (previously characterized) which serve as proton energy degraders  between energy positions.
At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor --- after end-of-beam (EoB), decay radiation emitted from these activated stainless steel foils may be used to develop radiochromic film (Gafchromic EBT3), revealing the spatial profile of the beam entering and exiting the stack.
After loading all targets in the stack, the lid of the target box is sealed in place, using an inset o-ring to create a water-tight seal, and the box is lowered through a hot cell into the beamline, where it sits electrically isolated.
The specifications of the target stack design for this work is presented in \autoref{tab:stack_table}.

This target stack was assembled and irradiated at the Isotope Production Facility (IPF) at the Los Alamos National Laboratory (LANL), using the LANSCE linear accelerator. 
The stack was irradiated for approximately 2 hours with a nominal current of 1 mA, using a 50 \micro s pulse at a frequency of 2 Hz, for an anticipated integral current of 205.9 nAh.
The beam current, measured using an inductive pickup, remained stable under these conditions for the duration of the irradiation, with the exception of approximately 70\,s of downtime, which occurred approximately 3 minutes into irradiation.
The proton beam incident upon the stack's Inconel beam entrance window had an average energy of 100 MeV determined via time-of-flight, with an approximately Gaussian energy distribution width of 0.1 MeV --- this energy profile was used for all later analysis.
At the end of the irradiation, the target stack was withdrawn from the beamline into the IPF hot cell, where it was disassembled and the activated foils removed using robotic manipulators.
The activated foils were cleaned of all surface contamination, and transported to a counting lab for gamma spectroscopy, which started approximately 7 hours following end-of-beam.

@@

For consistency, a single detector was used in this measurement, an ORTEC GEM Series (model \#GEM10P4-70)  High-Purity Germanium (HPGe) detector.
The detector is a mechanically-cooled coaxial p-type HPGe with a 1 mm aluminum window, and a 49.2 mm diameter, 27.9 mm long crystal.
Samples were counted at fixed positions ranging 4.5--83.5  cm (5\% maximum permissible dead-time) from the front face of the detector, with a series of standard calibration sources used to determine energy, efficiency, and pileup calibrations for each position.
The foils were counted  for a period of 2 weeks following EoB, to accurately quantify all induced activities,  with dead time never exceeding 5\%.
An example gamma-ray spectrum collected in such a fashion is shown in \autoref{fig:gspec}.
For all spectra collected, net peak areas were fitted using the gamma spectroscopy analysis code UNISAMPO \cite{Aarnio2001}, which has been shown to perform best in comparisons with other common analysis codes \cite{Jackman2014}.
%, due to its peak fitting algorithms and incorporation  of calibrated detector-specific peak shapes and widths.

@@

The \enquote{variance minimization} techniques  described by  Graves \etal\  have been employed here to further reduce the uncertainty in proton energy assignments     \cite{Graves2016}.
This method is based on the assumption that the independent measurements of proton fluence from the 5 monitor reactions used in this work should all be consistent at each energy position.
If the monitor reaction cross sections and MCNP6-modeled energy distributions are both accurate, then any disagreement in the  observed proton fluences is due to a systematic uncertainty in the stack design, namely, the areal densities of the stack components \cite{Graves2016,Marus2015}. 
This disagreement is minor at the front of the stack, and gets progressively worse as the beam is degraded, due to the compounded effect of systematic uncertainties in stack areal densities.
